---
title: "52414 - lab 1"
author: "52414"
date: "4/4/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

# *Lab 1: Basic Data Wrangling*  
<br/><br/>  
  

**Contents**:  

* Q0) [Submission Instructions](#submission-instructions)  
* Q1) [Data Preparation and Manipulation](#data-preparation-and-manipulation)      
* Q2) [Analysis of Daily New Corona Cases and Deaths](#analysis-of-daily-new-corona-cases-and-deaths)    
* Q3) [Preparing and Analyzing the World Bank Data](#preparing-and-analyzing-the-world-bank-data)
* Q4) [Joining the Datasets](#joining-the-datasets)  
* Q5) [Open Question](#open-question)

<br/><br/>
  
  
### Submission Instructions  
  
This lab will be submitted in pairs using GitHub (if you don't have a pair, please contact us).  
Please follow the steps in the  [GitHub-Classroom Lab 1](https://classroom.github.com/g/oSZNtHq4) to create your group's Lab 1 repository.  
**Important: your team's name must be `FamilyName1_Name1_and_FamilyName2_Name2`**.  
You can collaborate with your partner using the git environment; You can either make commits straight to master, or create individual branches (recommended). However, once done, be sure to merge your branches to master - you will be graded using the most recent master version - your last push and merge before the deadline.   
**Please do not open/review other peoples' repositories - we will be notified by GitHub if you do.**

Your final push should include this Rmd file (with your answers) together with the html file that is outputted automatically by knitr when you knit the Rmd. Anything else will be disregarded. In addition, please adhere to the following file format:    
`Lab_2_FamilyName1_Name1_and_FamilyName2_Name2.Rmd/html`      


<br/><br/>
  
The only allowed libraries are the following (**please do not add your own**):
```{r, include=FALSE}
library('tidyverse')
library(data.table)
```  
<br/><br/>

## A Deeper Dive Into John's Hopkins Corona Database         
    
The John's Hopkins Novel Corona Virus (COVID-19) epidemiological data is compiled by the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) from various sources. <br>
The dataset contains data since 22nd of January 2020. For the data and more information about it, please visit [here](https://data.humdata.org/dataset/novel-coronavirus-2019-ncov-cases).    
  
In this lab you will pick up where we left in lecture 2 and analyze the Corona cases and deaths data.  

### Q1
### Data Preparation and Manipulation   
(25 points)  

1. We first prepare and aggregate the data.   

a. First, load the `Corona Confirmed Cases Narrow`, the `Corona Confirmed Deaths Narrow`, and the `Corona Confirmed Recovered Narrow` datasets directly from the John's Hopkins website.  
The type of the `Date` variable should be date type. (2 pts)      
b. Create new data-frames named `cases.agg`, `deaths.agg`, and `recovered.agg` which aggregate the `sum` of Corona cases, deaths, and recovered respectively over the different countries' provinces. To do this, aggregate `Value` using only the country and date features, ignoring all other features (similarly to what has been shown in `lecture 2`).  
To achieve the aggregation use the `aggregate` function. In addition, order the data-frame first by Country and then by Date (increasing order). The columns of each of the two resulting data-frames should be `Country.Region, Date, Value`. (5pts)   
c. Repeat (b) using `tidyverse` and the pipe. Show that the outputs from the two methods are the same. (5pts)  
d. Using the last day of March as a reference, create a single stacked bar-plot that visualizes the top 10 countries in terms of their Corona cases, and their respected Corona deaths and recovered cases stacked on top of the current sick people in three different colors (each stack should add up to total cases). Make sure that the first stack shows the number of confirmed Corona sick people (`sick = cases - deaths - recovered`). Each stacked bar should represent a country. Please use the barplot base R function to create this plot. I recommend everyone to read the help documentation for barplot ; go over the expected input, and the different parameters. What is the biggest issue with the information presented in this plot? (13pts)

   
  
**Solution:**  
```{r}
## 1.a
corona_conf <- read.csv(url("https://data.humdata.org/hxlproxy/data/download/time_series_covid19_confirmed_global_narrow.csv?dest=data_edit&filter01=explode&explode-header-att01=date&explode-value-att01=value&filter02=rename&rename-oldtag02=%23affected%2Bdate&rename-newtag02=%23date&rename-header02=Date&filter03=rename&rename-oldtag03=%23affected%2Bvalue&rename-newtag03=%23affected%2Binfected%2Bvalue%2Bnum&rename-header03=Value&filter04=clean&clean-date-tags04=%23date&filter05=sort&sort-tags05=%23date&sort-reverse05=on&filter06=sort&sort-tags06=%23country%2Bname%2C%23adm1%2Bname&tagger-match-all=on&tagger-default-tag=%23affected%2Blabel&tagger-01-header=province%2Fstate&tagger-01-tag=%23adm1%2Bname&tagger-02-header=country%2Fregion&tagger-02-tag=%23country%2Bname&tagger-03-header=lat&tagger-03-tag=%23geo%2Blat&tagger-04-header=long&tagger-04-tag=%23geo%2Blon&header-row=1&url=https%3A%2F%2Fraw.githubusercontent.com%2FCSSEGISandData%2FCOVID-19%2Fmaster%2Fcsse_covid_19_data%2Fcsse_covid_19_time_series%2Ftime_series_covid19_confirmed_global.csv"), comment.char="#")
corona_death <- read.csv(url("http://data.humdata.org/hxlproxy/data/download/time_series_covid19_deaths_global_narrow.csv?dest=data_edit&filter01=explode&explode-header-att01=date&explode-value-att01=value&filter02=rename&rename-oldtag02=%23affected%2Bdate&rename-newtag02=%23date&rename-header02=Date&filter03=rename&rename-oldtag03=%23affected%2Bvalue&rename-newtag03=%23affected%2Binfected%2Bvalue%2Bnum&rename-header03=Value&filter04=clean&clean-date-tags04=%23date&filter05=sort&sort-tags05=%23date&sort-reverse05=on&filter06=sort&sort-tags06=%23country%2Bname%2C%23adm1%2Bname&tagger-match-all=on&tagger-default-tag=%23affected%2Blabel&tagger-01-header=province%2Fstate&tagger-01-tag=%23adm1%2Bname&tagger-02-header=country%2Fregion&tagger-02-tag=%23country%2Bname&tagger-03-header=lat&tagger-03-tag=%23geo%2Blat&tagger-04-header=long&tagger-04-tag=%23geo%2Blon&header-row=1&url=https%3A%2F%2Fraw.githubusercontent.com%2FCSSEGISandData%2FCOVID-19%2Fmaster%2Fcsse_covid_19_data%2Fcsse_covid_19_time_series%2Ftime_series_covid19_deaths_global.csv"), comment.char="#")
corona_rec <- read.csv(url("https://data.humdata.org/hxlproxy/data/download/time_series_covid19_recovered_global_narrow.csv?dest=data_edit&filter01=merge&merge-url01=https%3A%2F%2Fdocs.google.com%2Fspreadsheets%2Fd%2Fe%2F2PACX-1vTglKQRXpkKSErDiWG6ycqEth32MY0reMuVGhaslImLjfuLU0EUgyyu2e-3vKDArjqGX7dXEBV8FJ4f%2Fpub%3Fgid%3D1326629740%26single%3Dtrue%26output%3Dcsv&merge-keys01=%23country%2Bname&merge-tags01=%23country%2Bcode%2C%23region%2Bmain%2Bcode%2C%23region%2Bsub%2Bcode%2C%23region%2Bintermediate%2Bcode&filter02=merge&merge-url02=https%3A%2F%2Fdocs.google.com%2Fspreadsheets%2Fd%2Fe%2F2PACX-1vTglKQRXpkKSErDiWG6ycqEth32MY0reMuVGhaslImLjfuLU0EUgyyu2e-3vKDArjqGX7dXEBV8FJ4f%2Fpub%3Fgid%3D398158223%26single%3Dtrue%26output%3Dcsv&merge-keys02=%23adm1%2Bname&merge-tags02=%23country%2Bcode%2C%23region%2Bmain%2Bcode%2C%23region%2Bsub%2Bcode%2C%23region%2Bintermediate%2Bcode&merge-replace02=on&merge-overwrite02=on&filter03=explode&explode-header-att03=date&explode-value-att03=value&filter04=rename&rename-oldtag04=%23affected%2Bdate&rename-newtag04=%23date&rename-header04=Date&filter05=rename&rename-oldtag05=%23affected%2Bvalue&rename-newtag05=%23affected%2Binfected%2Bvalue%2Bnum&rename-header05=Value&filter06=clean&clean-date-tags06=%23date&filter07=sort&sort-tags07=%23date&sort-reverse07=on&filter08=sort&sort-tags08=%23country%2Bname%2C%23adm1%2Bname&tagger-match-all=on&tagger-default-tag=%23affected%2Blabel&tagger-01-header=province%2Fstate&tagger-01-tag=%23adm1%2Bname&tagger-02-header=country%2Fregion&tagger-02-tag=%23country%2Bname&tagger-03-header=lat&tagger-03-tag=%23geo%2Blat&tagger-04-header=long&tagger-04-tag=%23geo%2Blon&header-row=1&url=https%3A%2F%2Fraw.githubusercontent.com%2FCSSEGISandData%2FCOVID-19%2Fmaster%2Fcsse_covid_19_data%2Fcsse_covid_19_time_series%2Ftime_series_covid19_recovered_global.csv"), comment.char="#")
corona_conf$Date <- as.Date(corona_conf$Date)
corona_rec$Date <- as.Date(corona_rec$Date)
corona_death$Date <- as.Date(corona_death$Date)

## 1.b
# all the data affregated
cases.agg <- aggregate(Value ~ Country.Region + Date, data = corona_conf, FUN = sum)
deaths.agg <- aggregate(Value ~ Country.Region + Date, data = corona_death, FUN = sum)
recovered.agg <- aggregate(Value ~ Country.Region + Date, data = corona_rec, FUN = sum)

## 1.c
#all the data affregated with pipe
cases.agg.pipe <- corona_conf %>% group_by(Country.Region, Date) %>% summarise(Value = sum(Value))
deaths.agg.pipe <- corona_death %>% group_by(Country.Region, Date) %>% summarise(Value = sum(Value))
recovered.agg.pipe <- corona_rec %>% group_by(Country.Region, Date) %>% summarise(Value = sum(Value))

## 1.d
#all the data on 31/03/20
march <- data.frame(Country.Region = cases.agg$Country.Region[which(cases.agg$Date == "2020-03-31")],Value= cases.agg$Value[which(cases.agg$Date == "2020-03-31")],Date = cases.agg$Date[which(cases.agg$Date == "2020-03-31")])
march_max_v <- march[order(march$Value, decreasing = TRUE),] # data ordered by cases
temp_m <- c(head(march_max_v$Country.Region,10)) #top 10 index number by cases
#all the data about the top 10 countries by cases
all <- data.frame(Country.Region = march$Country.Region[temp_m], value.c = cases.agg$Value[which(cases.agg$Date == "2020-03-31")][temp_m], value.d = deaths.agg$Value[which(deaths.agg$Date == "2020-03-31")][temp_m], value.r = recovered.agg$Value[which(recovered.agg$Date == "2020-03-31")][temp_m])
#a data frame that has all the sick,deaths and recovered of the top 10 countries
temp_char <- data.frame(name = rep(c("sick","deaths","recovered"), each = 10), country = rep(march$Country.Region[temp_m],3),values = c(all$value.c - all$value.d - all$value.r, all$value.d, all$value.r))

colors = c("yellow","red","black") # a vector of the colors of the graph
matr <- matrix(c(temp_char$values),nrow = 3,ncol = 10,byrow = TRUE) # all the values in a matrix
country <- as.vector(march$Country.Region[temp_m]) #a vector of the top 10 countries
kind <- c( "Sick", "Recovered","Deaths") # a vector of the kinds of variables 

#a bar plot of the sick, recovered and deaths of the top 10 contries at the end of march

barplot(matr/1000, main = "Corona", names.arg = country,las = 2,  col = colors,ylim = c(0,200),ylab = "Value (Thousands)")
legend("topright", kind , col = colors , bty = "n", pch=20 , pt.cex = 3, cex = 0.8, horiz = FALSE, inset = c(0,0))
mtext("Country",side=1,col="black",line=4)

```


# 1.d
#The main issue with this way of showing the information, is that the countries that are shown in the final ggplot is of the top 10 countries with confirmed cases. But not all of those countries are the top ten of deaths and recovered. That means that the ggplot is not aggregated with the other categories.

<br/><br/>  

### Q2
### Analysis of Daily New Corona Cases and Deaths  
20 points

The two datasets (Corona Cases and Deaths) register the value of cases and deaths, respectively, as a cumulative sum for each day. In this question we would like to understand the daily differences between consecutive days.     

a. Add a new column named `Diff` to both the `cases.agg` and the `deaths.agg` data-frames. This new column should register the daily `Value` difference for each country. In other words, the `Diff` column shows how many new cases/deaths each country incurs every day. Hint - diff must be per country. (7pts)  
b. Find the top 10 instances of country and date combinations with the greatest absolute number of new daily Corona cases and deaths (separately). Print the result in a descriptive format. (5pts)  
c. In one figure, plot Italy's new daily Corona cases AND deaths as a function of Date. Choose the plot type you think that makes the most sense. (3pts) 
d. Plot the same graph as in (c), but this time plot the number of new cases on the logarithm scale. What can we learn? (5pts)  
 

**Solution:** 
```{r}
## 2.a
#added a column of the diffrence betwin today and the day befor to the cases
cases.agg <- cases.agg[order(cases.agg$Country.Region, decreasing = FALSE),]
cases.agg <- cases.agg %>%
    group_by(Country.Region) %>%
    mutate(Diff = Value - lag(Value))
#added a column of the diffrence betwin today and the day befor to the deaths
deaths.agg <- deaths.agg[order(deaths.agg$Country.Region, decreasing = FALSE),]
deaths.agg <- deaths.agg %>%
    group_by(Country.Region) %>%
    mutate(Diff = Value - lag(Value))

## 2.b
#top 10 cases and deaths
cases.agg <- cases.agg[order(cases.agg$Diff, decreasing = TRUE),]
deaths.agg <- deaths.agg[order(deaths.agg$Diff, decreasing = TRUE),]
max_cases_d <- data.frame(Country.Region = head(cases.agg$Country.Region,10), Date = head(cases.agg$Date,10), Diff = head(cases.agg$Diff,10)) 
max_death_d <- data.frame(Country.Region = head(deaths.agg$Country.Region,10), Date = head(deaths.agg$Date,10), Diff = head(deaths.agg$Diff,10)) 
max_cases_d
max_death_d

## 2.c
#all the data about the cases in italy
italy_cases <- data.frame(Cases = cases.agg$Diff[which(cases.agg$Country.Region == "Italy")],Date = cases.agg$Date[which(cases.agg$Country.Region == "Italy")] )
#all the data about the deaths in italy
italy <- data.frame( Date = deaths.agg$Date[which(deaths.agg$Country.Region == "Italy")], Cases = cases.agg$Diff[which(cases.agg$Country.Region == "Italy")], Deaths = deaths.agg$Diff[which(deaths.agg$Country.Region == "Italy")])
italy$Date <- as.Date(italy$Date)
italy <- italy[order(italy$Date,decreasing = FALSE),]

#a plot that shows the increase in cases and deaths by date in italy
par(mar=c(4,4,2,3.5))
plot(x = italy$Date, y = italy$Cases , main = "Italy", xlab="", ylab="",pch=16, yaxt="n",xaxt = "n",ylim = c(0,7000),type="l",xlim = c(min(as.Date(italy$Date)),max(as.Date(italy$Date))))
axis(2, ylim=c(0,7500),col="black",las=1)
mtext("Cases",side=2,line=3)
par(new=TRUE)
plot(x = italy$Date, y = italy$Deaths,axes = "False", pch = 16, col = "red",ylim = c(0,1000),xlab = "",ylab = "",type="l",xlim = c(min(as.Date(italy$Date)),max(as.Date(italy$Date))))
mtext("Deaths",side=4,col="red",line=2) 
axis(4, ylim=c(0,1000), col="red",col.axis="red",las=2)
mtext("Date",side=1,col="black",line=3)
axis(1, italy$Date, format(italy$Date, "%b %d"), cex.axis = 1)
legend("topleft",legend=c("Cases","Deaths"),bty = "n",
  text.col=c("black","red"),pch=c(16,15),col=c("black","red"))


## 2.d
#a scatter plot that shows the increase in cases by date in italy on Logarithmic Scale
plot(x = italy$Date, y = log(italy$Cases) , main = "Italy Logarithmic Scale", xlab = "Date", ylab = "Cases",ylim = c(0,10),type = "o",xaxt="n", pch = 16)
axis(1, italy$Date, format(italy$Date, "%b %d"), cex.axis = 1)
```

# 2.d
# At first there was an exponential increase and from the middel of march you can see that the graph became constant. we can learn that once Italy started to take things in to there hands and dill with the corona things started to get better ad insted of an exponential increase there is a constant increase.

<br/><br/>


### Q3
### Preparing and Analyzing the World Bank Data   
25 points

a. Rename the columns of `eco_data`: `country,S_country,feature,feature_code,Y2018V,Y2019V`. (2pts)  
b. Create a new `eco` data-frame whose dimensions are $N \times 11$, where `N` is the number of countries. The first column should include the names of the countries in `eco_data.`   
The rest of the columns should be the features with their respective values in `eco_data` for each country from 2018. Print the head of the new data-frame.(8pts).   
In other words, in this question you are asked to create a new eco data-frame whose dimensions are $N \times 11$, where N is the number of countries (the first column should include the names of the countries in `eco_data`).
The rest of the columns should be the features with their respective values in eco_data for each country from 2018. Print the head of the new data-frame. You may delete rows with NA as a country value only.  
c. Select and rename the following columns: `country` as country, `GDP(US currency)` as GDP, `Population ages 65 and above (% of total population)` as pop65, `Population in the largest city (% of urban population)` as pop_city_ratio, `Population, total` as pop_total columns .  (2pts) 
d. Show a table of the five countries with the highest per capita GDP in 2018.     
Next (considering all countries), plot the % of population over 65 vs. log of GDP per capita in 2018, after excluding the 10% countries with the lowest GDP per capita. Using `lm` and `abline`, add a regression line to the plot. What is your conclusion? (13 pts)  
  
  
  
**Solution:** 
  
```{r}
## 3.a
eco_data <- read.csv(url("https://raw.githubusercontent.com/DataScienceHU/DataAnalysisR_2020/master/data/economic_data.csv"))

#all eco data organized
eco_data <- eco_data %>% 
  rename(country = `ï..Country.Name`, S_country = Country.Code, feature = Series.Name, feature_code = Series.Code ,Y2018V = X2018..YR2018. ,Y2019V = X2019..YR2019.)

##3.b

eco <- head(eco_data,-5) # eco without the emty lines (last 5 lines)
eco <- dcast(eco,country~feature,value.var= "Y2018V") # eco data with the 2018 data
head(eco)

##3.c
#eco with only the columns that we need
eco <- select(eco, country, `GDP (current US$)`, `Population ages 65 and above (% of total population)`, `Population in the largest city (% of urban population)`, `Population, total`)

#rename the columns
eco <- eco %>% 
    rename(country = country, GDP = `GDP (current US$)`, pop65 = `Population ages 65 and above (% of total population)`, pop_city_ratio = `Population in the largest city (% of urban population)`, pop_total_columns = `Population, total`)

##3.d
GDP_per_ca <- as.numeric(eco$GDP)/as.numeric(eco$pop_total_columns) #Calculation of GDP per capita
eco_t <- eco #temporary eco that we can make changes on
eco_t$GDP_per_ca <- GDP_per_ca #added a column of the GDP per capita
max_GDP <- eco_t[order(eco_t$GDP_per_ca,decreasing = TRUE),] # eco ordered by the GDP per capeta
head(max_GDP,5)

GDP_per_ca_90 <- head(max_GDP,round(length(GDP_per_ca) * 0.9)) #the top 90% of countries by the GDP per capeta
x <- as.numeric(GDP_per_ca_90$pop65) #the population over 65
y <- as.numeric(GDP_per_ca_90$GDP_per_ca) # GDP per capeta
# a plot of the population over 65 vs the GDP per capeta
plot(x = x , y = log(y) , xlim = c(0,27), main = "Pop Over 65 Vs GDP", xlab = "GDP Per Capeta", ylab = "Population Over 65") 
regression_line <- lm(log(y) ~ x) #regression line
abline(regression_line, col = "red")

```



# 3.d
# you can see that there is a very strong correlation between the population over 65 and the GDP per capita

<br/><br/>  


### Q4
### Joining the Datasets   
20 points

a. Join the `deaths.agg`, `cases.agg`, and `recovered.agg` into one data-frame called `corona`.(5pts)
b. Join the `corona` and `eco` data-frames in a way that will keep the most information regarding the data (but not full join).   
Make sure that no essential data is thrown away (show this). (3pts)
c. Create new columns of normalized `cases`, `deaths`, and `recovered` so they will show the number of cases per 100,000 people for each country.   
Using the last day of March as a reference, create a single stacked bar plot that visualizes the top 10 countries in terms of normalized Corona cases, and their respected normalized Corona deaths and recovered, as done in Q1.   
how is it different from the graph before normalization? (5pts)
d. Using the last day of March as a reference, create a scatter-plot of normalized deaths and cases vs. `pop65`. Limit the plot to show only countries with 15% or more of `pop65`.   
In addition, color the outliers( pop65>24, norm100K_deaths>15) in that plot in red and add to the plot their country names (7pts)
  
  
**Solution:**   
```{r}
## 4.a
# all data back to order
cases.agg <- cases.agg[order(cases.agg$Country.Region),]
cases.agg <- cases.agg[order(cases.agg$Date),]
deaths.agg <- deaths.agg[order(deaths.agg$Country.Region),]
deaths.agg <- deaths.agg[order(deaths.agg$Date),]
recovered.agg <- recovered.agg[order(recovered.agg$Country.Region),]
recovered.agg <- recovered.agg[order(recovered.agg$Date),]

#a data of the corona  data in one nice data frame 
corona <- left_join(by = c("Date","Country.Region"),x = cases.agg,y = deaths.agg) %>% rename(country = "Country.Region", Cases.Value = "Value.x", Deaths.Value = "Value.y",Cases.Diff = "Diff.x", Deaths.Diff = "Diff.y")
corona$Recovered.Value <- recovered.agg$Value

#4.b
#a data of the corona and eco data in one nice data frame 
missing_countries <- anti_join(x = corona,y = eco,"country") #missing countries that are in corona and not in eco
missing_countries <- aggregate(Cases.Value ~ country,missing_countries,FUN = sum)
missing_countries2 <- anti_join(eco,corona,"country") #missing countries that are in eco and not in corona 
eco$country <- as.character(eco$country)

#changed the names in eco to be the same as in corona
eco$country[eco$country == "Bahamas, The"] <- "Bahamas"
eco$country[eco$country == "Brunei Darussalam"] <- "Brunei"
eco$country[eco$country == "Congo, Dem. Rep."] <- "Congo (Kinshasa)"
eco$country[eco$country == "Congo, Rep."] <- "Congo (Brazzaville)"
eco$country[eco$country == "Czech Republic"] <- "Czechia"
eco$country[eco$country == "Egypt, Arab Rep."] <- "Egypt"
eco$country[eco$country == "Gambia, The"] <- "Gambia"
eco$country[eco$country == "	Iran, Islamic Rep."] <- "Iran"
eco$country[eco$country == "Korea, Rep."] <- "Korea"
eco$country[eco$country == "Kyrgyz Republic"] <- "Kyrgyzstan"
eco$country[eco$country == "Lao PDR"] <- "Laos"
eco$country[eco$country == "Russian Federation"] <- "Russia"
eco$country[eco$country == "St. Kitts and Nevis"] <- "Saint Kitts and Nevis"
eco$country[eco$country == "St. Lucia"] <- "Saint Lucia"
eco$country[eco$country == "St. Vincent and the Grenadines"] <- "Saint Vincent and the Grenadines"
eco$country[eco$country == "Slovak Republic"] <- "Slovakia"
eco$country[eco$country == "Syrian Arab Republic"] <- "Syria"
eco$country[eco$country == "United States"] <- "US"
eco$country[eco$country == "Venezuela, RB"] <- "Venezuela"
eco$country[eco$country == "Yemen, Rep."] <- "Yemen"
eco$country[eco$country == "Sub-Saharan Africa"] <- "Western Sahara"

full_table <- right_join(x = corona,y = eco,"country") # a data frame of all the corona data and eco data

#4.c
#a data frame of all the corona data and eco data on 31/03/20
corona_march <- data.frame(country = corona$country[which(corona$Date == "2020-03-31")],  Cases.Value =  corona$Cases.Value[which(corona$Date == "2020-03-31")], Cases.Diff = corona$Cases.Diff[which(corona$Date == "2020-03-31")], Deaths.Value = corona$Deaths.Value[which(corona$Date == "2020-03-31")], Deaths.Diff = corona$Deaths.Diff[which(corona$Date == "2020-03-31")], Recovered.Value = corona$Recovered.Value[which(corona$Date == "2020-03-31")])
#a data frame of all the corona and  eco data on 31/03/20
full_march_table <- right_join(x = corona_march,y = eco,"country")

#added a per 100,000 column for cases,deaths and recovered to my data frame
full_march_table$Cases_per_100000 <- as.numeric(full_march_table$Cases.Value) / (as.numeric(full_march_table$pop_total_columns) / 100000) 
full_march_table$Deaths_per_100000 <- as.numeric(full_march_table$Deaths.Value) / (as.numeric(full_march_table$pop_total_columns) / 100000)
full_march_table$Recovered_per_100000 <- as.numeric(full_march_table$Recovered.Value) / (as.numeric(full_march_table$pop_total_columns) / 100000)

march_max_corona <- full_march_table[order(full_march_table$Cases_per_100000, decreasing = TRUE),] # data ordered by cases per 100,000
temp_m_per <- as.numeric(c(head(rownames(march_max_corona),10))) # a numeric vector of the index top 10 countries by per 100,000 cases
# all the per 100,000 data in one data frame
all_per <- data.frame(country = full_march_table$country[temp_m_per], value.c_per = full_march_table$Cases_per_100000[temp_m_per], value.d_per = full_march_table$Deaths_per_100000[temp_m_per], value.r_per = full_march_table$Recovered_per_100000[temp_m_per])

#a data frame that has all the sick,deaths and recovered of the top 10 countries
temp_char_per <- data.frame(name = rep(c("sick","deaths","recovered"), each = 10), country = rep(full_march_table$country[temp_m_per],3),values = c(all_per$value.c_per - all_per$value.d_per - all_per$value.r_per, all_per$value.d_per, all_per$value.r_per))

#all the values in a matrix
matr_per <- matrix(c(temp_char_per$values),nrow = 3,ncol = 10,byrow = TRUE) #all the values in a matrix
country_per <- as.vector(full_march_table$country[temp_m_per]) #a vector of the top 10 countries

#a bar plot of the sick, recovered and deaths of the top 10 contries at the end of march
barplot(matr_per, main = "Corona per 100,000", names.arg = country_per,las = 2, xlab = "Country", ylab = "Values", col = colors,ylim = c(0,800))
legend("topright", kind , col = colors , bty = "n", pch=20 , pt.cex = 2, cex = 0.8, horiz = FALSE, inset = c(0.05, 0.05))

#4.d
# the top 85% countries by populations over 65
march_table_15 <- data.frame(country = full_march_table$country[which(as.numeric(full_march_table$pop65) >= 15)], Cases_per_100000 = full_march_table$Cases_per_100000[which(as.numeric(full_march_table$pop65) >= 15)],  Deaths_per_100000 = full_march_table$Deaths_per_100000[which(as.numeric(full_march_table$pop65) >= 15)], pop65 = full_march_table$pop65[which(as.numeric(full_march_table$pop65) >= 15)] )


x <- full_march_table$pop65[which(as.numeric(full_march_table$pop65) >15)] #a vector of the top 85% pop over 65
y_case <- march_table_15$Cases_per_100000 #a vector of the top 85% cases by pop over 65
y_death <- march_table_15$Deaths_per_100000 #a vector of the top 85% deaths by pop over 65
country_15 <- march_table_15$country ##a vector of the top 85% countries by pop over 65
label <- ifelse(as.numeric(x) >24 | y_death > 15, as.character(country_15),"") #all the countries that have more than 24% over 65 population or the death per 100,000 is over 15

#a plot of the death and cases per 100,000 vs population over 65 with the outliers( pop65>24, norm100K_deaths>15) in red and the country name by the dots
plot(x, y_case ,pch = ifelse(as.numeric(x) >24, 15,0),col = ifelse(as.numeric(x) > 24,"red", "black"),main = "Death & Case Vs Pop Over 65", xlab = "Death & Case", ylab = "Pop 65")
points(x, y_death, pch = ifelse(as.numeric(x) >24 | y_death > 15, 16,1),col = ifelse(as.numeric(x) > 24 | y_death > 15,"red", "black"))
with(march_table_15,text( y_death ~ x, labels = label))
```


# 4.c
# we can see that once you look at the normalized facts the smaller countries are shown on the graph compared to the non normalized graph that the larger countries are shown. Spain and Italy are average countries and are shown on both graphs because the corona hit the strongest in the world

<br/><br/>  




### Q5
### Open Question
10 points
  
Write an interesting research question regarding the Corona outbreak and then follow the steps to answer it using tables and plots. You can use the loaded datasets or any other dataset you find as long as you add the data file to your `lab1` repository so it can be loaded directly from a `url` (e.g. the World Bank). This question will be graded based on creativity, originality, and the novelty of the analysis.   
  
**Solution:**   

```{r}
#data of how many tests per each confirmed case
test_vs_case <- data.frame(Country = number_of_covid_19_tests_per_confirmed_case$Entity, Date = number_of_covid_19_tests_per_confirmed_case$Date, Test_Per_Confirmed = number_of_covid_19_tests_per_confirmed_case$`Number of tests per confirmed case`)
test_vs_case$Country <- as.character(test_vs_case$Country)
test_vs_case$Country[test_vs_case$Country == "United States"] <- "US"
#data of how many cases are done in each day for every country
tests <- data.frame(Country = full_list_total_tests_for_covid_19$Entity, Date = full_list_total_tests_for_covid_19$Date, Tests = full_list_total_tests_for_covid_19$`Total tests`)
cases.agg <- cases.agg %>% 
  rename(Country = Country.Region, Diff_Cases = Diff, Cases = Value) #rename so the join will work
tests$Date <- as.Date(tests$Date)
test_vs_case$Date <- as.Date(test_vs_case$Date)
tests <- inner_join(tests,cases.agg,by = c("Country","Date")) # joining all the data into one data set
tests <- tests[order(tests$Country),]
tests <- tests %>%
  group_by(Country) %>%
  mutate(Diff_Tests = Tests - lag(Tests))
Total_pop <- data.frame(Country = eco$country, Population = eco$pop_total_columns) #a data with all the countries population
tests <- inner_join(tests,Total_pop,"Country")

#normalizing the data to 100K 
tests$Tests_Per_100k <- ((as.numeric(as.character(tests$Diff_Tests)))*100000) /(as.numeric(as.character(tests$Population)))
tests$Cases_Per_100k <- ((as.numeric(as.character(tests$Cases)))*100000) /(as.numeric(as.character(tests$Population)))
tests <- inner_join(tests,test_vs_case,by = c("Country","Date"))
tests$Date <- as.Date(tests$Date)

#data of two dates with two weeks apart
april_02 <- data.frame(Country = tests$Country[which(tests$Date == "2020-04-02")], Cases_Per_100k = tests$Cases_Per_100k[which(tests$Date == "2020-04-02")], Test_Per_Confirmed = tests$Test_Per_Confirmed[which(tests$Date == "2020-04-02")]) 
april_15 <-data.frame(Country = tests$Country[which(tests$Date == "2020-04-15")], Cases_Per_100k = tests$Cases_Per_100k[which(tests$Date == "2020-04-15")], Test_Per_Confirmed = tests$Test_Per_Confirmed[which(tests$Date == "2020-04-15")])
both <- inner_join(april_02,april_15,"Country")

#data and matrix for the bar plot
df = data.frame(Country = c(both$Country), Cases_April_15 = c(both$Cases_Per_100k.y/10), con_test_April_02 = c(both$Test_Per_Confirmed.x))
value_matrix <- matrix(,nrow = 2,ncol = 41)
value_matrix[1,] = df$Cases_April_15
value_matrix[2,] = df$con_test_April_02
value_matrix_round <- round(value_matrix,digits = 2)
#a bar plot that shows both variables for each country
mp <- barplot(value_matrix, names.arg = df$Country, beside = TRUE, las = 2, col = c("yellow", "red"),main = "Approved Tests Vs Cases",xlab = "Countries",ylim = c(0,60) )
legend("topright",c("Tests Approved In April 02", "Cases In April 15"),col = c( "red" ,"yellow"),bty = "n", pch=20 , pt.cex = 4, cex = 0.8, horiz = FALSE,inset = 0)
text(mp,value_matrix, labels = value_matrix_round, pos = 3, cex = .75) 

plot_countries <- c("Italy","Israel","Austria","Japan","United Kingdom")
plot_color <- c("black","blue","orange","red","yellow")

#plot of 5 random countries the tests per 100K 
plot(tests$Date[which(tests$Country == "Italy")],tests$Tests_Per_100k[which(tests$Country == "Italy")], ylim = c(0,200),pch = 20, main = "Tests Per 100K",xaxt = "n",xlab = "", ylab = "Cases",type = "o")
par(new = TRUE)
plot(tests$Date[which(tests$Country == "Israel")],tests$Tests_Per_100k[which(tests$Country == "Israel")],col = "blue", ylim = c(0,200),pch = 20,axes = FALSE ,type = "o", xlab = "", ylab = "")
par(new = TRUE)
plot(tests$Date[which(tests$Country == "Austria")],tests$Tests_Per_100k[which(tests$Country == "Austria")],col = "orange", ylim = c(0,200),pch = 20,axes = FALSE ,type = "o", xlab = "", ylab = "")
par(new = TRUE)
plot(tests$Date[which(tests$Country == "Japan")],tests$Tests_Per_100k[which(tests$Country == "Japan")],col = "red", ylim = c(0,200),pch = 20,axes = FALSE ,type = "o", xlab = "", ylab = "")
par(new = TRUE)
mtext("Date",side=1,col="black",line=3)
axis(1, tests$Date[which(tests$Country == "United Kingdom")], format(tests$Date[which(tests$Country == "United Kingdom")], "%b %d"), cex.axis = 1)
legend("topleft", plot_countries , col = plot_color , bty = "n", pch=20 , pt.cex = 3, cex = 0.8, horiz = FALSE, inset = c(0,0))

#plot of 5 random countries that shows a Logarithmic Scale of cases per 100K 
plot(tests$Date[which(tests$Country == "Italy")],log(tests$Cases_Per_100k[which(tests$Country == "Italy")]), ylim = c(0,6), pch = 20, main = "Cases Per 100K", xaxt = "n",xlab = "", ylab = "Cases",type = "o")
par(new = TRUE)
plot(tests$Date[which(tests$Country == "Italy")],log(tests$Cases_Per_100k[which(tests$Country == "Italy")]),col = "black", ylim = c(0,6), pch = 20,axes = FALSE ,type = "o", xlab = "", ylab = "")
par(new = TRUE)
plot(tests$Date[which(tests$Country == "Israel")],log(tests$Cases_Per_100k[which(tests$Country == "Israel")]),col = "blue", ylim = c(0,6), pch = 20,axes = FALSE ,type = "o", xlab = "", ylab = "")
par(new = TRUE)
plot(tests$Date[which(tests$Country == "Austria")],log(tests$Cases_Per_100k[which(tests$Country == "Austria")]),col = "orange", ylim = c(0,6), pch = 20,axes = FALSE ,type = "o", xlab = "", ylab = "")
par(new = TRUE)
plot(tests$Date[which(tests$Country == "Japan")],log(tests$Cases_Per_100k[which(tests$Country == "Japan")]),col = "red", ylim = c(0,6), pch = 20,axes = FALSE ,type = "o", xlab = "", ylab = "")
par(new = TRUE)
plot(tests$Date[which(tests$Country == "United Kingdom")],log(tests$Cases_Per_100k[which(tests$Country == "United Kingdom")]),col = "yellow", ylim = c(0,6), pch = 20,axes = FALSE ,type = "o", xlab = "", ylab = "")
mtext("Date",side=1,col="black",line=3)
axis(1, tests$Date[which(tests$Country == "United Kingdom")], format(tests$Date[which(tests$Country == "United Kingdom")], "%b %d"), cex.axis = 1)
legend("topleft", plot_countries , col = plot_color , bty = "n", pch=20 , pt.cex = 3, cex = 0.8, horiz = FALSE, inset = c(0,0))
```

<br/><br/>  